Conformance Checking Using Embeddings and its Applicability in the Internet of Things

### 2022-01-28 paper.tex

\section{Embedding Based Conformance Checking}
\label{sec:method}
Many conformance checking algorithms, notably the classical ones outlined above, aim to measure conformance between a log and a model.
While this is the fundamental question to be answered, this comparison between two different types of objects greatly complicates the task.
However, by re-framing the problem it  can be greatly simplified.
Instead of formulating the problem in terms of model-to-log comparison, one can formulate it as a comparison problem of a real log to a model log.
Such a reformulation of the problem is possible since all (reasonably defined) models are able to generate sample traces, and can therefore produce a model log.
Then the whole problem is re-framed as log-to-log comparisons, i.e. determining similarities between objects of the same type, for which a broad range of techniques already exist.
This has the additional advantage that it is model-independent since no limitations apart from the ability to generate traces are put on the type of model.
\begin{figure}
	\centering
	\includegraphics[width=.8\textwidth]{figures/structure}
	\caption{The general structure of the approach}
	\label{fig:structure}
\end{figure}

Three different techniques were outlined in the focus paper \cite{PBWe20}, all following a similar structure outlined in \cref{fig:structure}.
First, a model log is produced from the model by using appropriate replay semantics for the model.
Then, given these two logs, one real and one produced by replaying the model, it is possible to train the \emph{act2vec} \cite{KBWe18} embedding.
Based on these embeddings a dissimilarity function needs to be defined to compare different traces, however, the \emph{act2vec} embedding only yields distance measurements between activities and not between traces.
Here the differences between the three techniques come into play.
The first two, the Word Mover's Distance \cite{KSKW15} and Iterative Constrained Transfers, can directly utilize the distances between activities to determine a distance between two collections of activities.
The last one, \emph{trace2vec}, further needs to train a trace embedding based on activity embeddings, and can then perform direct comparisons in the trace embedding space.
All three methods however lead to a reasonable dissimilarity metric between traces.
Using one of these metrics a dissimilarity matrix can be built up describing the pairwise dissimilarities between the real log and the model log.

@@

\subsubsection{Iterative Constrained Transfers}
\begin{algorithm}
	\caption{ACT}\label{alg:act}
	\begin{algorithmic}
		\State $k:=$ number of edges considered per activity
		\State $p_i,q_i := $ normalized count of activity $i$ in $p$ or $q$ respectively
		\State $c(i,j) :=$ Distance between embeddings of activity $i$ and $j$
		\State $h_p,h_q :=$ amount of different activities in trace $p$ and trace $q$ 
		\Function{$F_{ICT}$}{$p,q,c,k$}
		\For{$i=1...h_p$}
		\State $s \gets \arg \min_k(c(i,[1,...,h_q]))$ \Comment{Find $k$ activities with the smallest cost}
		\While{$l < k$}
		\State $r \gets \min(p_i,q_{s(l)}))$ \Comment{Edge constraint, move as much as possible but no more than allowed}
		\State $p_i \gets p_i - r$ \Comment{Move weight}
		\State $t \gets t + rc(i,j)$ \Comment{Update cost}
		\State $l \gets l+1$
		\EndWhile
		\If{$p_i \neq 0$} \Comment{Move any remaining weight in $p_i$ to $q_{s(k)}$}
		\State $t \gets t + p_ic(i,s(k))$ \Comment{Move rest to $q_{s(k)}$}
		\EndIf
		\EndFor
		\State \Return $t$
		\EndFunction
	\end{algorithmic}
\end{algorithm}
The EMD and variants of it are very attractive since they are both very general and exhibit good accuracy for both search and classification.
A major drawback of such methods however is the cubic complexity in the size of the input distributions, which makes it infeasible for larger problems.
In the case of conformance checking the size of the input, distribution is determined by the number of activities in the log, for which good scalability is desired.
There are a number of techniques that aim to reduce the complexity of EMD-based approaches, that obtain an approximation of the EMD by relaxing some of the constraints.  
One example of such a method is the Relaxed Word Mover's distance which drops the second (inflow) constraint, thus greatly simplifying the solution.
While this makes it easier to calculate the solution, the RWMD only gives a lower bound on the WMD.
Another such method is the Iterative Constrained Transfers \cite{AtMi18} algorithm that has linear complexity in the size of the input distribution.
Instead of just dropping the inflow constraint it is instead replaced by an edge capacity constraint.
The main benefit of this is that it avoids the formulation as a linear program and is instead solvable using a much less powerful, but much quicker, direct algorithm.

>>>>

\section{Embedding Based Conformance Checking}
\label{sec:method}
Many conformance checking algorithms, notably the classical ones outlined above, aim to measure conformance between a log and a model.
While this is the fundamental question to be answered, this comparison between two different types of objects greatly complicates the task.
However, by re-framing the problem it  can be greatly simplified.
Instead of formulating the problem in terms of model-to-log comparison, one can formulate it as a comparison problem of a real log to a model log.
Such a reformulation of the problem is possible since all (reasonably defined) models are able to generate sample traces.
Then the key part is determining similarities between objects of the same type, for which a broad range of techniques already exist.
This has the additional advantage that it is model-independent since no limitations apart from the ability to generate traces are put on the type of model.
\begin{figure}
	\centering
	\includegraphics[width=.8\textwidth]{figures/structure}
	\caption{The general structure of the approach}
	\label{fig:structure}
\end{figure}

Three different techniques were outlined in the focus paper \cite{PBWe20}, all following a similar structure outlined in \cref{fig:structure}.
First, a model log is produced from the model by using appropriate replay semantics for the model.
Then, given these two logs, it is possible to train the \emph{act2vec} \cite{KBWe18} embedding\footnote{Note that only the activity labels were used to train the embedding here. It is possible to include further attributes, but this is more complicated}.
Then a dissimilarity function needs to be defined, to compare to bags of \emph{act2vec} embeddings to each other.
%Based on these embeddings a dissimilarity function needs to be defined to compare different traces, however, the \emph{act2vec} embedding only yields distance measurements between activities and not between traces.
Here the differences between the three techniques come into play.
The first two, the Word Mover's Distance \cite{KSKW15} and Iterative Constrained Transfers, can directly utilize the distances between activities to determine a distance between two collections of activities.
The last one, \emph{trace2vec}, further needs to train a trace embedding based on activity embeddings, and can then perform direct comparisons in the trace embedding space.
Using one of these metrics a dissimilarity matrix can be built up describing the pairwise dissimilarities between the real log and the model log.

@@

\subsubsection{Iterative Constrained Transfers}
\begin{algorithm}
	\caption{ACT}\label{alg:act}
	\begin{algorithmic}
		\State $k:=$ number of edges considered per activity
		\State $p_i,q_i := $ normalized count of activity $i$ in $p$ or $q$ respectively
		\State $c(i,j) :=$ Distance between embeddings of activity $i$ and $j$
		\State $h_p,h_q :=$ amount of different activities in trace $p$ and trace $q$ 
		\Function{$F_{ICT}$}{$p,q,c,k$}
		\For{$i=1...h_p$}
		\State $s \gets \arg \min_k(c(i,[1,...,h_q]))$ \Comment{Find $k$ activities with the smallest cost}
		\While{$l < k$}
		\State $r \gets \min(p_i,q_{s(l)}))$ \Comment{Edge constraint}
		\State $p_i \gets p_i - r$ \Comment{Move weight}
		\State $t \gets t + rc(i,j)$ \Comment{Update cost}
		\State $l \gets l+1$
		\EndWhile
		\If{$p_i \neq 0$} \Comment{Move any remaining weight in $p_i$ to $q_{s(k)}$}
		\State $t \gets t + p_ic(i,s(k))$ \Comment{Move rest to $q_{s(k)}$}
		\EndIf
		\EndFor
		\State \Return $t$
		\EndFunction
	\end{algorithmic}
\end{algorithm}
The EMD and variants of it are very attractive since they are both very general and exhibit good accuracy for both search and classification.
A major drawback of such methods however is the cubic complexity in the size of the input distributions, which makes it infeasible for larger problems.
In the case of conformance checking the size of the input, distribution is determined by the number of activities in the log, for which good scalability is desired.
There are a number of techniques that aim to reduce the complexity of EMD-based approaches, that obtain an approximation of the EMD by relaxing some of the constraints.  
One example of such a method is the Relaxed Word Mover's distance which drops the second (inflow) constraint, thus greatly simplifying the solution.
While this makes it easier to calculate the solution, the RWMD only gives a lower bound on the WMD.
Another such method is the Iterative Constrained Transfers \cite{AtMi18} algorithm that has linear complexity in the size of the input distribution.
Instead of just dropping the inflow constraint it is instead replaced by an edge capacity constraint.
The main benefit of this is that it avoids the formulation as a linear program and is instead solvable using a much less powerful, but much quicker, direct algorithm.
